---
title: "03_homework_193DS"
author: "Ian Del Tredici"
date: "03-05-2026"
format: html
---

[Github repo](https://github.com/iandeltredici/ENVS-193DS_homework-03.git)

```{r}
#| label: read in packages and data
#| message: false
library(tidyverse) #load in packages
library(here)
library(janitor)
library(readxl)
library(ggeffects)
salinity <- read.csv( #read in pickleweed data using here function
  here("data","salinity-pickleweed.csv"))
reading_data <- read.csv( #read in personal data using here function
  here("data", "reading_data_envs-193DS.csv")
)
```

# Part 2 Problems
## Problem 1. Slough soil salinity

###a. An appropriate test
Both a Spearman's correlation and a Pearson's correlation test could determine the relationship strength between salinity, a continuous predictor variable, and California pickleweed biomass, a continuous response variable. Pearson's correlation is a parametric test, meaning is assumes normality and uses the actual measurements of continuous data to show the strength and direction of a linear relationship. Meanwhile, Spearman's correlation uses ranks instead of the actual values to determine a monotonic relationship between variables, making it more suitable for non-normal distributions.

### b. Creae a visualization
```{r}
#| label: visualization salinity x biomass
#| message: false
salinity_lm <-lm(
  pickleweed ~ salinity_mS_cm,
  data = salinity)
summary(salinity_lm)
par(mfrow = c(2, 2))
plot(salinity_lm)
salinity_preds <- ggpredict(
  salinity_lm,
  terms = "salinity_mS_cm"
)
ggplot(data = salinity,
        aes(x = salinity_mS_cm,
            y = pickleweed)) +
  geom_point(size = 4,
             stroke = 1,
             fill = "forestgreen",
             shape = 21) +
  geom_ribbon(data = salinity_preds,
              aes(x = x,
                  y = predicted,
                  ymin = conf.low,
                  ymax = conf.high),
              alpha = 0.1) +
  geom_line(data = salinity_preds,
            aes(x = x,
                y = predicted)) +
  labs(x = "Salinity mS/cm",
       y = "Pickleweed biomass (g)",
       title = "As salinity increases, pickleweed biomass increases") +
    theme_minimal()
```

###c. Check your assumptions and run your test
#### Running the test
```{r}
#| label: run test
cor.test(salinity$pickleweed, salinity$salinity_mS_cm,
         method = "pearson")
```

#### Check assumptions
```{r}
#| label: checking assumptions
#| message: false
par(mfrow = c(2, 2))
plot(salinity_lm)
```

The assumptions checked for a Pearson correlation include normality by using a QQ plot to assess the linear relationship between the theoretical and sample quartiles, homoscedasticity with a residuals vs fitted and squared residuals vs fitted plot to check for constant variance amongs errors and an outlier plot. The QQ plot showed a linear relationship, indicating normality, similar to the constant variance plots that display fairly normal data with only a slight deviation at the tails. There were no perceived outliers.